### 2021-03-29 outputs/exploratory.Rmd
add some of the exploratory work

---
title: "data exploration"
author: "Amy"
date: "29/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

```{r}
isrdata <- read.dta13("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta")
```

```{r}
expdata <- isrdata %>%
  filter(use == 1)
```

```{r}
colSums(is.na(expdata))
```

```{r}
expdata %>%
  select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7) %>%
  unique() %>%
  summarize(white = sum(t1_race_1, na.rm = TRUE)/n(),
            black = sum(t1_race_2, na.rm = TRUE)/n(),
            hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
            asian = sum(t1_race_4, na.rm = TRUE)/n(),
            americanindian = sum(t1_race_5, na.rm = TRUE)/n(),
            middleeastern = sum(t1_race_6, na.rm = TRUE)/n(),
            other = sum(t1_race_7, na.rm = TRUE)/n(),
            total = n())
```

```{r}
expdata %>%
  group_by(treatment) %>%
  summarize(white = sum(s1_race_1, na.rm = TRUE)/n(),
            black = sum(s1_race_2, na.rm = TRUE)/n(),
            hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
            asian = sum(s1_race_4, na.rm = TRUE)/n(),
            americanindian = sum(s1_race_5, na.rm = TRUE)/n(),
            middleeastern = sum(s1_race_6, na.rm = TRUE)/n(),
            other = sum(s1_race_7, na.rm = TRUE)/n(),
            total = n())
```
```{r}
expdata %>%
  group_by(treatment) %>%
  summarize(meanage = mean(s1_age, na.rm = TRUE))
```
```{r}
expdata %>%
  group_by(treatment) %>%
  summarise(prestudentsimilarity = mean(s1_sim, na.rm = TRUE),
            prestudentISR = mean(s1_tsr),
            poststudentsimilarity = mean(s2_sim, na.rm = TRUE),
            poststudentISR = mean(s2_tsr, na.rm = TRUE),
            postinstructorsimilarity = mean(t2_sim1, na.rm = TRUE),
            postinstructorISR = mean(t2_tsr, na.rm = TRUE),
            standardgrade = mean(grade_sd, na.rm = TRUE))
```