### 2021-04-09 outputs/exploratory.Rmd
add models

---
title: "data exploration"
author: "Amy"
date: "29/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(readstata13)
library(arsenal)
library(Hmisc)
library(MASS)
library(kableExtra)
library(apaTables)
```

```{r}
codebook <- read_csv(here("inputs/data/codebook.csv"))
```

```{r, warning=FALSE}
isrdata <- read.dta13(here("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta"))
```

```{r}
expdata <- isrdata %>%
  filter(use == 1)
```

```{r}
summary(tableby(treatment ~  s1_sim + s1_tsr + s2_sim + s2_tsr + t2_sim1 + t2_tsr + grade + std_grade + t2_finalexam + f17_enrolled + ir_f16_gpa + s_female,
                data = isrdata)) %>%
  kable()
```

```{r}
rcorr(as.matrix(isrdata %>% 
                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa)))
```

```{r}
apa.cor.table(isrdata %>% 
                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa))
```

```{r}
colSums(is.na(expdata))
```

```{r}
expdata %>%
  select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7) %>%
  unique() %>%
  summarize(white = sum(t1_race_1, na.rm = TRUE)/n(),
            black = sum(t1_race_2, na.rm = TRUE)/n(),
            hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
            asian = sum(t1_race_4, na.rm = TRUE)/n(),
            americanindian = sum(t1_race_5, na.rm = TRUE)/n(),
            middleeastern = sum(t1_race_6, na.rm = TRUE)/n(),
            other = sum(t1_race_7, na.rm = TRUE)/n(),
            total = n())
```

```{r}
expdata %>%
  group_by(treatment) %>%
  summarize(white = sum(s1_race_1, na.rm = TRUE)/n(),
            black = sum(s1_race_2, na.rm = TRUE)/n(),
            hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
            asian = sum(s1_race_4, na.rm = TRUE)/n(),
            americanindian = sum(s1_race_5, na.rm = TRUE)/n(),
            middleeastern = sum(s1_race_6, na.rm = TRUE)/n(),
            other = sum(s1_race_7, na.rm = TRUE)/n(),
            total = n())
```
```{r}
expdata %>%
  group_by(treatment) %>%
  summarize(meanage = mean(s1_age, na.rm = TRUE))
```
```{r}
expdata %>%
  group_by(treatment) %>%
  summarise(prestudentsimilarity = mean(s1_sim, na.rm = TRUE),
            prestudentISR = mean(s1_tsr),
            poststudentsimilarity = mean(s2_sim, na.rm = TRUE),
            poststudentISR = mean(s2_tsr, na.rm = TRUE),
            postinstructorsimilarity = mean(t2_sim1, na.rm = TRUE),
            postinstructorISR = mean(t2_tsr, na.rm = TRUE),
            standardgrade = mean(grade_sd, na.rm = TRUE))
```

```{r}
summary(lm(data = isrdata, s1_sim ~ treatment))
```
```{r}
summary(lm(data = isrdata, s2_sim ~ treatment))
```
```{r}
summary(lm(data = isrdata, s2_tsr ~ treatment + s1_tsr))
```
```{r}
summary(polr(formula = as.factor(t2_sim1) ~ treatment, data = isrdata, Hess = TRUE))
```

```{r}
summary(lm(data = isrdata, s2_tsr ~ treatment))
```

```{r}
summary(lm(data = isrdata, grade ~ treatment + s1_female + ir_f16_gpa))
```

```{r}
summary(lm(data = isrdata %>% filter(obj_exam == 1), t2_finalexam ~ treatment + s1_female + ir_f16_gpa))
```

```{r}
summary(glm(data = isrdata, f17_enrolled ~ treatment + s1_female + ir_f16_gpa, family = "binomial"))
```