### 2021-04-27 outputs/paper/paper.Rmd
actual last edits.

---
title: "The inefficacy of superficial similarities for improving instructor-student relationships"
author: "Amy Farrow"
date: "April 26th, 2021"
header-includes:
  \usepackage{dcolumn}
  \usepackage{float}
  \usepackage{rotating}
  \usepackage{pdflscape}
  \usepackage{booktabs}
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Replicating ‘Taking It to the Next Level’"
abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationships. The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity, or instructor-student relationship. The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no measurable effects on any of the student outcome measures, including grades and reenrollment. These results are consistent with the original paper.\\par \\textbf{Keywords:} replication study, instructor-student relationships, college, similarity, college persistence"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in an RProject
library(Hmisc) # for correlation matrix
library(readstata13) # for stata files
library(finalfit) # for summary table
library(stargazer) # for model table
library(MASS) # for ordinal logistic model
library(cowplot) # for putting plots side by side
library(ltm) # for checking scale validity
library(DiagrammeR) # for causal diagram
library(tidymodels) # for train test
library(performance) # for model performance evaluation

# NOTE: script 01_ must have been run already

apply_if <- function(mat, p, f) {
  # Fill NA with FALSE
  p[is.na(p)] <- FALSE
  mat[p] <- f(mat[p])
  mat
}

apaCorr <- function(mat, corrtype = "pearson") {
  matCorr <- mat
  if (class(matCorr) != "rcorr") {
    matCorr <- rcorr(mat, type = corrtype)
  }

  # Add one star for each p < 0.05, 0.01, 0.001
  stars <- apply_if(round(matCorr$r, 2), matCorr$P < 0.05, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.01, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.001, function(x) paste0(x, "*"))
  # Put - on diagonal and blank on upper diagonal
  stars[upper.tri(stars, diag = T)] <- "-"
  stars[upper.tri(stars, diag = F)] <- ""
  n <- length(stars[1,])
  #colnames(stars) <- 1:n
  # Remove _ and convert to title case
  #row.names(stars) <- tools::toTitleCase(sapply(row.names(stars), gsub, pattern="_", replacement = " "))
  # Add index number to row names
  #row.names(stars) <- paste(paste0(1:n,"."), row.names(stars))
  stars
}
```

# Introduction

Despite the growing importance of college degrees for career success, college completion rates are a continuing problem. The causes are complex: factors that affect student persistence include student characteristics, organizational factors, peer environment, and student experiences. Interventions to improve student persistence can target any of these aspects.

@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college student persistence and performance. In this field experiment, they evaluated the links between perceived similarity, instructor-student relationships (ISRs), and student success. Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive impact on measurable student outcomes. The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not. Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.

This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors. First, @citeoriginal's methods are discussed, and the measures and scales are evaluated. Second, demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated. Third, @citeoriginal's models are replicated, and additional models are explored. Linear, logistic, and ordinal logistic models are replicated to predict outcomes including ISR perception, grade, and persistence.  Further replicated exploratory models consider the relationship between similarity and instructor-student relationship, as well as between instructor-student relationship and student outcomes. Additional models use different controls--specifically, variables indicating shared racial and gender self-identification.

These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence. These conclusions stand when matching instructor-student traits are considered and when vulnerable student groups are evaluated separately from the wider student population. The lack of significant effects reflects the complicated multi-stage task of increasing perceived similarity, to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.

## Literature review

College is perceived to be a meritocratic tool for social mobility and career success [@college]. Unfortunately, retention is a large problem in contemporary American colleges, and many students begin degrees without completing them; six-year completion rates for full-time first-time students range from 51% to 86%, depending on school [@finishline]. Even when controlling for pre-college test scores and initial enrollments, completion rate disparities exist based on parental education, socio-economic status, and race/ethnicity [@finishline]. Disparities in college completion lead to further entrenchment of long-standing inequalities [@finishline]. Thus, measures to help students persist in college completion are desirable to reduce wasted resources and to increase societal equity.

Many causes, and corresponding solutions, have been theorized for the 'college completion crisis'. In his examination of persistence research, @Reason_2009 considers student precollege characteristics (sociodemographic traits, academic preparation and performance, and dispositions), organizational factors (structural-demographic characteristics and organizational behavior dimensions), student peer environment (campus racial and academic climates), and individual student experiences (curricular, classroom, and out-of-class experiences). Such reviews indicate the sheer complexity involved. Implementing an effective intervention is difficult due to complicated causes of attrition, embedded social inequalities, and expenses [@Tinto_2006].

Interventions targeting non-academic factors are popular. For example, using a randomized controlled trial, @citecasemanagement implemented a case management intervention, where social workers helped students with financial assistance, course selection, finding childcare, and accessing social services. The case management intervention significantly improved persistence rates [@citecasemanagement]. First year programs that are separate from academic faculty are popular, but increasingly there is a focus on the classroom's importance for retention [@Tinto_2006]. @Tinto_1997 evaluated the impact of a college program designed for communal learning, which aims to engage students more than traditional programs. They found that students in the communal program participated in more academic activities, reported more positive views of the college and their role in it, and persisted at higher rates than the control [@Tinto_1997].

Especially at the elementary and secondary level, many interventions to improve student outcomes do so by targeting ISRs. @Roorda_Koomen_Spilt_Oort_2011 performed a meta-analysis of 99 studies on teacher-student relationships and student engagement and achievement. Effect sizes were larger for engagement than for achievement, and the effect sizes varied widely based on the specific measurements being used [@Roorda_Koomen_Spilt_Oort_2011]. They found stronger effects of teacher-student relationships in the upper secondary level, as compared to the lower elementary level [@Roorda_Koomen_Spilt_Oort_2011]. At the college level, @Tinto_2006 argues that the instructor is a key player in student retention.  @Creasey_Jarvis_Gadke_2009 found that students who had connected, nonthreatening relationships with their instructors had more positive feelings about their own abilities and expectations for the course.

ISRs are built on a range of social behaviors and experiences, but similarity's effect on ISRs is of particular interest for this study. This is a specific case of a wider field of study: how does similarity affect all types of relationships? In general, people are more likely to persist in building relationships with acquaintances that they share similarities with, including prejudices, behaviors, personality traits, attitudes, demographics, and activities [@Bahns_Crandall_Gillath_Preacher_2017]. In mentoring relationships, perceived deep similarity contributes to information sharing behavior and positive reception, which then is positively associated with mentee adjustment [@Zheng_Zheng_Wu_Yao_Wang_2021].

Other studies consider similarity and student performance, but do not directly address ISRs. One study of Taiwanese high school students explored the link between shared education and life values and student performance [@Lai_2015]. While they found a significant, but small, positive relationship between life values and student performance on analytic tests, education values gave mixed results depending on the natures of the education value and the analytic test in question [@Lai_2015]. At the college level, @Abrami_Mizener_1985's study found a small correlation between students' perceiving instructors' attitudes to be similar to their own and the grades of those students, but the correlation became insignificant when instructor fixed effects were considered. In the college context, the link between similarity and grades is not well-established.

Beyond the wide range of research that addresses similarity, ISRs, and student outcomes, 'Taking It to the Next Level' was directly inspired by two previous works in middle and high schools. @Gehlbach_Brinkworth_Harris_2012's study of middle-school students found that students' perceived similarity to their teachers positively correlated with improvements in teacher-student relationship, and that changes in teacher-student relationships were associated with changes in student perceived self-efficacy, but not scholastic performance. Using a randomized controlled trial and surveys, @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015 found that when high school students were presented with similarities they shared with their teachers, they subsequently perceived greater similarity with their teachers. When teachers were presented with similarities, they perceived better relationships with those students, and those students subsequently performed better academically. For students, however, awareness of similarities did not effectively improve their perception of the student-teacher relationship.

# Methodology

```{r, warning=FALSE, message=FALSE, echo=FALSE}
isrdata <- read.dta13(here("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta"))
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
expdata <- isrdata %>%
  filter(t1_consent == 1 # Only include if instructor consented to first survey
         & t1_complete == 1 # Only included if instructor completed first survey
         & t1_timer_consent_3 > 1 # Only include if instructor read consent
         & s1_consent == 1 # Only include if student consented to first survey
         & s1_timer_consent_3 > 1 # Only include if student read consent
         & s1_fullsurvey == 1 # Only include if the student completed the whole first survey
         & t2_complete == 1 # Only include if the instructor completed the whole second survey
         # Make sure that the survey taker looked at each page for more than three seconds
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & s1_timer_gtky1_3 > 10
         & s1_timer_gtky2_3 > 10
         & s1_timer_gtky3_3 > 10
         & s1_timerpre1_3 > 10
         & s1_timerpre2_3 > 10
         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
         & s1_timerpost1_3 > 10
         & s1_timerpost2_3 > 10
         & s1_timerdemo1_3 > 10
         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))
         & t2_timer > 10
         # Remove specific drop reasons
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         )

expdata <- expdata %>%
  mutate(s1_female = ifelse(s1_female == -99, NA, s1_female),
         t_female = ifelse(t_female == -99, NA, t_female)) %>% # Some NAs are shown as -99. Replace with NA.
  mutate(teacherid = as.factor(teacherid),
         f17_enrolled = as.factor(f17_enrolled),
         s1_female = as.factor(s1_female),
         t_female = as.factor(t_female))
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# The original coding of race contained errors -- ex. a professor who selected White and Asian in the survey,
# and was marked as Asian for most cases, but white for one random case.
# This remakes the variable t_race, for consistency

expdata <- expdata %>%
  mutate(t1_race_1 = replace_na(as.numeric(t1_race_1), 0),
         t1_race_2 = replace_na(as.numeric(t1_race_2), 0),
         t1_race_3 = replace_na(as.numeric(t1_race_3), 0),
         t1_race_4 = replace_na(as.numeric(t1_race_4), 0),
         t1_race_5 = replace_na(as.numeric(t1_race_5), 0),
         t1_race_6 = replace_na(as.numeric(t1_race_6), 0),
         t1_race_7 = replace_na(as.numeric(t1_race_7), 0)) %>%
  mutate(t_race = ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 > 1, 
                         "Mixed race/ethnicity", 
                         ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 < 1, "Unknown",
                                ifelse(t1_race_1 == 1, "White/Caucasian",
                                       ifelse(t1_race_2 == 1, "Black or African American",
                                              ifelse(t1_race_3 == 1, "Hispanic American or Latino/a",
                                                     ifelse(t1_race_4, "Asian or Pacific Islander",
                                                            ifelse(t1_race_5, "American Indian or Alaskan",
                                                                   ifelse(t1_race_6, "Middle Eastern",
                                                                          "Other Race")))))))))

expdata <- expdata %>%
  mutate(s1_race_1 = replace_na(as.numeric(s1_race_1), 0),
         s1_race_2 = replace_na(as.numeric(s1_race_2), 0),
         s1_race_3 = replace_na(as.numeric(s1_race_3), 0),
         s1_race_4 = replace_na(as.numeric(s1_race_4), 0),
         s1_race_5 = replace_na(as.numeric(s1_race_5), 0),
         s1_race_6 = replace_na(as.numeric(s1_race_6), 0),
         s1_race_7 = replace_na(as.numeric(s1_race_7), 0)) %>%
  mutate(s_race = ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 > 1,
                         "Mixed race/ethnicity", 
                         ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 < 1, "Unknown",
                                ifelse(s1_race_1 == 1, "White/Caucasian",
                                       ifelse(s1_race_2 == 1, "Black or African American",
                                              ifelse(s1_race_3 == 1, "Hispanic American or Latino/a",
                                                     ifelse(s1_race_4, "Asian or Pacific Islander",
                                                            ifelse(s1_race_5, "American Indian or Alaskan",
                                                                   ifelse(s1_race_6, "Middle Eastern",
                                                                          "Other Race"))))))))) %>%
  mutate(s_hisp_black = ifelse((s1_race_2 == 1 | s1_race_3 == 1), 1, 0))

# This creates a variable that is the final grade if the final was an objectively graded exam
# and otherwise NA
expdata <- expdata %>%
  mutate(t2_finalexamobj = ifelse(obj_exam == 1, t2_finalexam, NA))
```

@citeoriginal conducted a randomized controlled trial, with surveys, to assess the impact of awareness of instructor-student similarities on similarity perception, instructor-student relationship perception, grades, and reenrollment.

## Participants

The study took place in the 2017 spring term at a large Californian University. The study included 148 instructors and their 3,355 students. The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized. Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors. The initial sample of 148 instructors was a convenience sample: the study was advertised, interested instructors signed up and were enrolled if their course met the study requirements. 145 instructors consented and took the initial survey, resulting in an corresponding sample of 3,352 students. Once errors of administration, missing responses, and inadequate time spent on surveys were considered, the sample consisted of only 119 instructors and 2,273 students.

## Treatment and control

Participating students were randomly assigned to either treatment or control. At the beginning of the term, all participating students and instructors were given "get to know you" surveys. Using those responses, for each student in the treatment group, seven commonalities were identified between student and instructor (for example, perhaps both student and instructor binge-watch TV to relieve stress, or appreciate loyalty as the most important friend quality), and both student and instructor were informed of these commonalities. They completed a few questions about the similarities and were reminded of them through the term to ensure they were internalized. Students in the control group were informed about similarities they shared with students in another part of the country, and instructors were told nothing about these students.

## Procedures

At the beginning of the term, the instructors were surveyed, to collect potential similarities. Then, students were surveyed, to collect potential similarities. Immediately, students were either given the treatment or the placebo, and then given further survey questions that asked about similarity and ISR. Students were surveyed again at the end of the term, regarding similarity and ISR. Instructors were surveyed about similarity and ISR only at the end of the term.

## Measures

@citeoriginal identify key measures. Full descriptions of all measures are available in [Appendix A].

  1. Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5. These responses were averaged to create a *student perceived similarity scale*.
  2. Immediately after the treatment or placebo, students answered seven questions about their anticipated instructor-student relationship, on scales of 1 to 5. These responses were averaged to create a *student anticipated ISR scale*.
  3. At the end of the term, students answered the *student perceived similarity scale* questions again.
  4. At the end of the term, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create a *student perceived ISR scale*.
  5. At the end of the term, instructors answered only one question about *instructor perceived similarity* with the student, on a scale of 1 to 5.
  6. At the end of the term, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create an *instructor ISR perception* scale.
  7. Instructors were asked to report the student's *grade on their final exam*, paper, or project.
  8. The *grade that the student received in the course* was pulled from internal records.
  9. The student's course grade was *standardized* against other grades in the course.
  10. The student's *status as of Fall term 2017*, either enrolled or not, was pulled from internal records.

### Scale Reliability

A scale is externally reliable if it gives the same results across different testings [@Coolican_2014]. Test-retest reliability, where a scale is given to a group at different times, is a common way to assess this [@Coolican_2014]. In this data, there is a retest, but interactions between instructors and students may reasonably be expected to affect the second testing, rendering the data unuseful for assessing test-retest reliability.

A scale is internally reliable if it is consistent within itself [@Coolican_2014]. Cronbach's alpha was used to assess the reliability of the similarity and ISR scales. This statistic evaluates how much participants vary on individual items, compared to how they vary overall [@Coolican_2014]. Good reliability is indicated by an alpha between .75 and 1 [@Coolican_2014]. For the initial student perceived similarity scale, student anticipated ISR scale, end-of-term student  perceived similarity scale, end-of-term student perceived ISR scale, and end-of-term instructor perceived ISR scale, $\alpha$ = 0.892, 0.872, 0.91, 0.905, and 0.926, respectively, indicating that the scales are reliable.

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# Conduct tests for scale reliability
s1_sim_scale <- expdata %>%
  dplyr::select(s1_sim1, s1_sim2, s1_sim3, s1_sim4, s1_sim5, s1_sim6) %>%
  drop_na()
cronbach.alpha(s1_sim_scale)

s1_tsr_scale <- expdata %>%
  dplyr::select(s1_tsr1, s1_tsr2, s1_tsr3, s1_tsr4, s1_tsr5, s1_tsr6, s1_tsr7) %>%
  drop_na()
cronbach.alpha(s1_tsr_scale)

s2_sim_scale <- expdata %>%
  dplyr::select(s2_sim1, s2_sim2, s2_sim3, s2_sim4, s2_sim5, s2_sim6) %>%
  drop_na()
cronbach.alpha(s2_sim_scale)

s2_tsr_scale <- expdata %>%
  dplyr::select(s2_tsr1, s2_tsr2, s2_tsr3, s2_tsr4, s2_tsr5, s2_tsr6, s2_tsr7) %>%
  drop_na()
cronbach.alpha(s2_tsr_scale)

t2_tsr_scale <- expdata %>%
  dplyr::select(t2_tsr1, t2_tsr2, t2_tsr3, t2_tsr4, t2_tsr5, t2_tsr6, t2_tsr7) %>%
  drop_na()
cronbach.alpha(t2_tsr_scale)
```

### Scale Validity

A scale is valid if it measures what it is supposed to measure [@Coolican_2014]. In this case, the scales should measure similarity perception and instructor-student relationship (ISR) perception. @citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, and they were developed for use with sixth-, seventh-, and eight-grade American students and their teachers. Given the fact that these scales were developed for children, it is worth considering what the scales actually measure. To do this, we can look at the specific questions in the three distinct scales.

To assess similarity, students were asked about values, course goals, views on course content, general commonalities, personality, and overall similarity. By this standard, high similarity is when a student shares values, course goals, and views on course content, and the student and instructor have similar personalities, high commonality, and high general similarity. Questions ask about course-specific measures (content and goals), personal traits (values and personality), and non-specific traits (commonalities and general similarity), but do not ask about life experience, abilities, or approaches, amoungst other ways that two people can theoretically be similar.

To assess ISR, students were asked about enjoyment of learning, friendliness, encouragement, excitement, motivation, caring, and overall learning. By this standard, a high-quality ISR is one where a student enjoys learning from, is motivated by, and learns a great deal from a friendly, encouraging, caring instructor, who would be excited to see them in three years time. To assess ISR, instructors were asked about enjoyment of aiding learning, caring, frequency of encouragement, friendliness, excitement, motivation, and overall learning. By this standard, a high-quality ISR is one where an instructor enjoys helping and frequently encourages a friendly, caring student who is motivated and learns a great deal from this instructor, and the instructor would be excited to see the student in three years.

As @Brinkworth_McIntyre_Juraschek_Gehlbach_2018 argue, the instructor-student relationship is conceptualized, and therefore measured, in a variety of ways. Other constructions of ISR scales may measure negative aspects, respect, liking, responsiveness, engagement, trust, honesty, humour, interest, approachability, clarity, confidence, fairness, patience, and consistency [@Brinkworth_McIntyre_Juraschek_Gehlbach_2018]. @citeoriginal's ISR scale is measuring only one concept of an ISR, and that particular concept needs to be considered when analyzing the results of this experiment. Like most research about ISRs, @citeoriginal's scale originates in elementary and secondary schools, and the concept of "good ISR" may not translate easily to a post-secondary context.

# Data

This paper uses the R statistical language [@citeR] and `tidyverse` packages [@citetidyverse]. `here` was used to manage the RProject [@citehere]. Stata files were handled using `readstata13` [@citereadstata13]. The report was made using RMarkdown [@citermarkdown] and `bookdown` [@citebookdown]. Tables were made using `kableExtra` [@citekableExtra], `finalfit` [@citefinalfit], and `stargazer` [@citestargazer]. The correlation matrix was made using functions by @apacorrcode using the `Hmisc` package [@citeHmisc]. `MASS` provided additional models [@citeMASS]. Plots used `cowplot` [@citecowplot]. `ltm` was used to check scale validity [@citeltm]. `tidymodels` [@citetidymodels] and `performance` [@citeperformance] were used in assessing model validity and performance. Finally, the diagram was made using `DiagrammeR` [@citeDiagrammeR].

The data, in a .dta format, and Stata code for the @citeoriginal paper are available on the Inter-University Consortium for Political and Social Research [@Robinson_Scott_Gottfried_2020]. The dataset contains 36,838 observations of 653 variables. While there are many additional variables, including a wider range of survey questions targeting interest level, enjoyment, and confidence, this paper focuses on the covariates and key outcomes that are discussed in the original paper.

## Demographics

Most key demographic traits come from the initial surveys given to students and instructors. For gender, they were asked to identify as male, female, or other, and given the option to enter text. For race/ethnicity, they were allowed to check any number of a range of options ('White/Caucasian', 'Black or African American', 'Hispanic American or Latino/a', 'Asian or Pacific Islander', 'American Indian or Alaskan', 'Middle Eastern', and 'Other Race'). They were also asked if they were the first generation in their family to attend college (no or yes). Finally, they were asked to enter their age in years. The student's year of study and cumulative CGPA as of the previous term and the class size came from the college's internal records.

Table \@ref(tab:demotable1) shows the student covariates for the treatment and control samples. There are no significant differences between the treatment and control groups, as indicated by p-values which are all above 0.4.

```{r demotable1, echo=FALSE, warning=FALSE, message=FALSE}
# Make a summary table of nominal variables, by treatment group.
demotab <- expdata %>% 
  dplyr::select(treatment, 
                s1_female, 
                s_race, 
                s_firstgen,
                s1_age,
                ir_f16_gpa,
                year) %>%
  mutate(year = as.factor(year)) %>%
  mutate(treatment = ff_label(treatment, "Group"),
         s1_female = ff_label(s1_female, "Student gender"), 
         s1_age = ff_label(s1_age, "Student age"),
         s_race = ff_label(s_race, "Student race/ethnicity"), 
         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
         s_firstgen = ff_label(s_firstgen, "Student first-gen status"), 
         year = ff_label(year, "Year")) %>%
  summary_factorlist("treatment", 
                     c("s1_female", "s_race", "s_firstgen", "year", "s1_age", "ir_f16_gpa"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N(%) = ") %>%
  rename(N = "Total N", Missing = "Missing N")

# Replace variables that are encoded with their true meanings
demotab[2:3,4] <- c("Male", "Female")
demotab[13:14,4] <- c("No", "Yes")

# Put it all in a nice table.
demotab %>%
  knitr::kable(caption = "Student covariates for treatment and control groups",
               booktabs = TRUE, linesep = "",
               ) %>%
  column_spec(5, width = "5em") %>%
  column_spec(6, width = "5em") %>%
  column_spec(7, width = "5em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
                            stripe_index =c(2:3, 13:14, 20),
                            stripe_color = "#f6f0fb")
```

Table \@ref(tab:demotable2) shows the instructor covariates. Because instructors are counted multiple times, for each student in the treatment and control groups that is in their course, only the totals are shown. The same covariates, presented by treatment and control groups, are shown in [Appendix B].

```{r demotable2, echo=FALSE, warning=FALSE, message=FALSE}
# make a summary table of continuous variables
demotab2 <- expdata %>%
  dplyr::select(teacherid, t_age, t_firstgen, t_race, t_female, n_course) %>%
  unique() %>%
  mutate(teacherid = ifelse(as.numeric(teacherid) > 50, "0", "1")) %>%
  mutate(t_age = ff_label(t_age, "Instructor age"), 
         t_female = ff_label(t_female, "Instructor gender"),
         t_race = ff_label(t_race, "Instructor race/ethnicity"),
         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status"),
         n_course = ff_label(n_course, "Class size")) %>%
  summary_factorlist("teacherid",
                     c("t_female", "t_race", "t_firstgen", "t_age", "n_course"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     add_col_totals = TRUE,
                     col_totals_prefix = "N = ") %>%
  rename(N = "Total N", Missing = "Missing N")

demotab2[2:3,4] <- c("Male", "Female")
demotab2[10:11,4] <- c("No", "Yes")

# put in nice table
demotab2 %>%
  dplyr::select(-"0", -"1", -"p") %>%
  knitr::kable(caption = "Teacher covariates",
               booktabs = TRUE, linesep = "",
               col.names = c("", "N", "Missing", "", "")) %>%
  kableExtra::kable_styling(latex_options = c("striped"), 
                            stripe_index =c(2:3, 10:11, 13),
                            stripe_color = "#f6f0fb")
```

There are differences between the students and instructors. Notably, the student sample is 62.7% female, while the instructor sample is 78.2% female. The student sample is 19.8% White, while the instructor sample is 38.7% White. The student sample is 47% Hispanic, while the instructor sample is only 3.4%. Finally, the student sample is 9.8% multiracial, but the instructor sample is 52.1% multiracial. Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classified differently as a result. For this reason, it is very difficult to draw conclusions. Additionally, students are 43.5% first-generation college students, while only 25.2% of instructors are.

In light of the different percentages in student and instructor racial self-identification, Figure \@ref(fig:racegraph) shows the percentages of each group that selected a given racial identity, as well as the percentage that selected more than one of these categories. While the tables display race/ethnicity as a single variable (it is either a single identity, mixed race/ethnicity, or unknown), this graph, and the subsequent analysis, deal with multiple racial/ethnic identities that are either selected or not. On the graph, we can see that more students identified as Hispanic, while more instructors identified as White and selected multiple options. Viewing only the table, where race is shown as a one-dimensional variable, does not show the nuance of multiple racial/ethnic identities. Still, these categories are limiting, and do not necessarily truly reflect the social or cultural identities of study participants.

```{r racegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8,4), fig.cap="Bar chart illustrating racial identification proprtions of participants"}

# Summarise instructor race/ethnicity data
instructorrace <- expdata %>%
  dplyr::select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7, t_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(t1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(t1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(t1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(t1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(t1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(t1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(t1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(t_multi_race > 1, na.rm = TRUE)/n())

# summarise student race/ethnicity data
studentrace <- expdata %>%
  dplyr::select(id,
         s1_race_1, s1_race_2, s1_race_3, s1_race_4, s1_race_5, s1_race_6, s1_race_7, s_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(s1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(s1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(s1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(s1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(s1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(s1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(s1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(s_multi_race > 1, na.rm = TRUE)/n())

# bind the two summaries and present them in a bar chart
bind_cols(as_tibble(cbind(nms = names(instructorrace), t(instructorrace))),
          as_tibble(cbind(nms = names(studentrace), t(studentrace))))%>%
  dplyr::select(-nms...3) %>%
  pivot_longer(!nms...1, names_to = "group", values_to = "portion") %>%
  mutate(nms...1 = fct_relevel(nms...1, 
                               c('Hispanic American or Latino/a',
                                 'White/Caucasian',
                                 'Asian or Pacific Islander',
                                 'Middle Eastern',
                                 'Black or African American', 
                                 'American Indian or Alaskan',
                                 'Other Race',
                                 'Multiple')),
         group = fct_relevel(group, c("V2...4", "V2...2"))) %>%
  ggplot() +
  geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
  labs(x = "Self-identified race/ethnicity", y = "Proportion of sample", 
       fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
  scale_fill_manual(values = c("#c8b5f7", "#f9b5c9"), labels = c("Students", "Instructors")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
  theme_minimal()
```

Given the potential significance of class size for instructor-student relationships, Figure \@ref(fig:classsizegraph) shows the distribution of class sizes of instructors that participated in the study. Most class sizes are between 25 and 50 students, but there are some that are much larger.

```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8.5,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}

# Summarise course sizes from instructor POV
n_instructor <- expdata %>%
  dplyr::select(teacherid, n_course) %>%
  unique()

# Plot instructor POV course sizes
n_by_instructor <- n_instructor %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_instructor %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of classes") +
  theme_minimal()

# Summarise course sizes from a student POV
n_student <- expdata %>%
  dplyr::select(id, n_course) %>%
  unique()

# Plot student POV summary
n_by_student <- n_student %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_student %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of students in study") +
  theme_minimal()

# Combine the plots in a grid
combined <- plot_grid(n_by_instructor, n_by_student,
          ncol = 2)
title <- ggdraw() + 
  draw_label(
    "Some classes are very large (class size means marked)",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.1, 2))
```

## Outcomes of interest [^1]

[^1]: Interactive graphs and tables showing outcomes of interest are available [here](https://amycfarrow.shinyapps.io/takingittothenextlevel/).

Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values. There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception (p < 0.001 for initial perception, and p = 0.004 for end-of-term perception). Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 for intial perception and 3.5 versus 3.6 for end-of-term perception).

```{r summarytable, echo=FALSE, warning=FALSE, message=FALSE}
# Summarize key outcome variables by treatment group
sumtable <- expdata %>% 
  dplyr::select(treatment, 
                s1_sim,
                s1_tsr,
                s2_sim,
                s2_tsr,
                t2_sim1,
                t2_tsr, grade,
                std_grade,
                t2_finalexam,
                f17_enrolled) %>%
  mutate(t2_sim1 = as.factor(t2_sim1)) %>%
  summary_factorlist("treatment",
                     c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr",
                       "grade", "std_grade", "t2_finalexam", "f17_enrolled"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N(%) = ") %>%
  rename(N = "Total N", Missing = "Missing N")

# Rename variables for readability
sumtable[,1] <- c("", "Similarity", "ISR", " Similarity", " ISR", "Similarity ", "", "", "", "", "ISR ", "Course grade", "Stand. grade", "Final grade", "Peristence", "")
# replace nominal encodings with true values
sumtable[15,4] <- "No"
sumtable[16,4] <- "Yes"

# put in table.
sumtable %>%
  knitr::kable(caption = "Outcomes of interest for treatment and control groups",
               booktabs = TRUE, linesep = "",
               col.names = c("", "N", "Missing", "", "Control", "Treatment", "Total", "p")) %>%
  pack_rows("Initial student perception", 2, 3) %>%
  pack_rows("End-of-term student perception", 4, 5) %>%
  pack_rows("End-of-term instructor perception", 6, 11) %>%
  pack_rows("Student outcomes", 12, 15) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
                            stripe_index =c(2,4,6:10,12,14),
                            stripe_color = "#f6f0fb")
```

Figure \@ref(fig:continuousoutcomes) shows key continuous outcome distributions for both treatment and control. Corresponding to Table \@ref(tab:summarytable), the only visible differences in distributions are in the initial student similarity perception and end-of-term student similarity perception.

Interestingly, there is a large difference between how students and instructors rate perceived similarity and ISR. Students, on average, rate end-of-term similarity at 3.6, while instructors give an average rating of 2.7. Students give an average end-of-term ISR rating of 4.1, while instructors give an average of only 3.5. This could be due to more positive student perceptions, or due to students giving higher ratings for equivalent perceptions. Both groups rate ISR more highly than similarity, but it is difficult to compare the weight of these scores as the scales are not necessarily measuring in the same units.

```{r continuousoutcomes, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height= 10, fig.cap="Histograms illustrating continous and ordinal key outcomes"}
# Plot initial student similarity perception
s1_simplot <- expdata %>% 
  ggplot(aes(x = s1_sim, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Initial student similarity perception", y = "Frequency", fill = "Group") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot initial student ISR perception
s1_tsrplot <- expdata %>% 
  ggplot(aes(x = s1_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Student ISR anticipation", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot end-of-term student similarity perception
s2_simplot <- expdata %>% 
  ggplot(aes(x = s2_sim, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term student similarity perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# PLot end of term student ISR perception
s2_tsrplot <- expdata %>% 
  ggplot(aes(x = s2_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term student ISR perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot end of term instructor similarity perception
t2_sim1plot <- expdata %>% 
  ggplot(aes(x = t2_sim1, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term instructor similarity perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot end of term instructor ISR perception
t2_tsrplot <- expdata %>% 
  ggplot(aes(x = t2_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term instructor ISR perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot grade
gradeplot <- expdata %>% 
  ggplot(aes(x = grade, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Course grade", y = "Frequency") +
  scale_x_continuous(limits = c(-.1,4.5)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# plot standardized grades
std_gradeplot <- expdata %>% 
  ggplot(aes(x = std_grade, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Standardized course grade", y = "Frequency") +
  scale_x_continuous(limits = c(-4.5,3)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot final exam grades
t2_finalexamplot <- expdata %>% 
  ggplot(aes(x = t2_finalexam, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Final grade", y = "Frequency") +
  scale_x_continuous(limits = c(-.1,4.5)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Combine plots in a grid
combined <- plot_grid(s1_simplot + theme(legend.position = "none"), s1_tsrplot + theme(legend.position = "none"),
          s2_simplot + theme(legend.position = "none"), s2_tsrplot + theme(legend.position = "none"),
          t2_sim1plot + theme(legend.position = "none"), t2_tsrplot + theme(legend.position = "none"),
          std_gradeplot + theme(legend.position = "none"), t2_finalexamplot + theme(legend.position = "none"),
          ncol = 2)
legend <- get_legend(s1_simplot + 
                       guides(color = guide_legend(nrow = 1)) +
                       theme(legend.position = "bottom"))
title <- ggdraw() + 
  draw_label(
    "Most key outcomes unaffected by treatment",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.1, 2, .1))
```

Replicating @citeoriginal's study, Table \@ref(tab:correlationtable) displays a correlation matrix for the measures that @citeoriginal identified as key. The very strong correlations (course grade and the  standardized version, course grade and final grade, and course grade and objectively graded final grade) and some of the strong correlations (standardized course grade with final grade and objectively grade final grade) are entirely expected.

Other strong correlations are more interesting:  initial student similarity and instructor-student relationship perceptions, end-of-term student similarity and instructor-student relationship perceptions, and end-of-term instructor similarity and instructor-student relationship perceptions. At a given moment in time, a participant's assessment of similarity and ISR are highly related. This could indicate that the two perception scales may not be measuring psychologically distinct concepts, or that feelings of similarity and good relationships are strongly associated in the classroom.

Moderate correlations exist between initial and end-of-term student similarity perceptions as well as initial and end-of-term student ISR perceptions, indicating that student perceptions change somewhat but not entirely through the term. There are some moderate correlations between instructor perceptions of similarity or ISR and course grade, final grade, and objectively graded final grade, perhaps indicating that instructor opinions of students are linked to student performance in some way.

```{r correlationtable, echo=FALSE, warning=FALSE, message=FALSE}
# Make the summary stats part of the table
summary <- expdata %>% 
  dplyr::select(treatment, 
                s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, 
                grade, std_grade, t2_finalexam, t2_finalexamobj) %>%
  summary_factorlist("treatment", 
                     c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr",
                       "grade", "std_grade", "t2_finalexam", "t2_finalexamobj"),
                     p = FALSE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = FALSE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N (%) = ") %>%
  dplyr::select("Total N", "Missing N", "Total")

# Make the correlation matrix part of the table
matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr,
                                grade, std_grade, t2_finalexam, t2_finalexamobj))))

# Combine the summary and correlation tables
table <- summary %>%
  bind_cols(matrix) %>%
  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N")

# rename the variables for table readability
rownames(table) <- c("Similarity", "ISR", " Similarity", " ISR", "Similarity ", "ISR ", "Course", "Stand. course", "Final", "Obj. Final")

# Put in a table.
table[,1:11] %>%
  kable(caption = "Correlation matrix for continuous and ordinal key variables and outcomes",
        booktabs = TRUE,
        col.names = c("N", "Missing", "Mean (SD)", "Similarity", "ISR", "Similarity","ISR",
                      "Similarity", "ISR", "Course", "Stand. course")) %>%
  column_spec(1, width = "7em") %>%
  column_spec(4, width = "4em") %>%
  column_spec(5, width = "4em") %>%
  column_spec(6, width = "4em") %>%
  column_spec(7, width = "4em") %>%
  column_spec(8, width = "4em") %>%
  column_spec(9, width = "4em") %>%
  column_spec(10, width = "4em") %>%
  column_spec(11, width = "4em") %>%
  pack_rows("Initial student perception", 1, 2) %>%
  pack_rows("End-of-term student perception", 3, 4) %>%
  pack_rows("End-of-term instructor perception", 5, 6) %>%
  pack_rows("Grades", 7, 10) %>%
  add_header_above(c(" " = 4, "Initial student" = 2, "End-of-term student" = 2,
                     "End-of-term instructor" = 2, "Grades" = 2), 
                   align = "l") %>%
  kable_styling(latex_options = c("scale_down"))
```

## Missing data

There are no significant patterns in missing data. [Appendix C] shows the distributions of key variables relative to missing data in other key variables.

## Attrition

140 eligible instructors completed the first survey, and 2,808 of their students completed the first survey. 21 of those instructors did not complete the second survey. 2,808 students completed the first survey (1,396 in the treatment group, 1,412 in the control group), but 755 of those students did not complete the second survey (364 in the treatment group, 391 in the control group). The attrition in the treatment and control groups was comparable (26.1% in the treatment group, 27.7% in the control group).

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors participated? 148
isrdata %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# How many students could theoretically participate, based on their instructor participating? 3355
isrdata %>%
  filter(!is.na(teacherid)) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors consented to the first survey? 145
isrdata %>%
  filter(t1_consent == 1) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# How many students does that correspond to? 3352
isrdata %>%
  filter(t1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# How many of those students consented to survey 1? 3229
isrdata %>%
  filter(t1_consent == 1, s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# So total participants in the study: 145 instructors and 3229 students.
# Now reasons for removal from analysis.
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# If we consider non-eligibility, how many instructors are left? 140
isrdata %>%
  filter(t1_consent == 1 & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students") %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# And students? 3112
isrdata %>%
  filter(t1_consent == 1 & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# Check for instructors completing the first survey. 140
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t1_complete == 1) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# The number of students that corresponds to. 3112
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Then, how many students completed the first survey? 2808
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

## Treatment: 1396
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & treatment == "Treatment") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

## Control 1412
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1 
         & treatment == "Control") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors completed the second survey? 119
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t1_complete == 1
         & t2_complete == 1) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# How many students does that correspond to? 2701
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Students that completed the second survey: 2053
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & s2_complete) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Treatment 1032
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & s2_complete
         & treatment == "Treatment") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Control 1021
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & s2_complete 
         & treatment == "Control") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many teachers spent enough time on the survey? 119.
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t1_complete == 1
         & t2_complete == 1
         & t1_timer_consent_3 > 1
         & s1_timer_consent_3 > 1
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & t2_timer > 10) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# Students that corresponds to: 2560
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & t1_timer_consent_3 > 1
         & s1_timer_consent_3 > 1
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & t2_timer > 10) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Students that spent enough time on the surveys: 2273
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & t1_timer_consent_3 > 1
         & s1_timer_consent_3 > 1
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & t2_timer > 10
         & s1_timer_gtky1_3 > 10
         & s1_timer_gtky2_3 > 10
         & s1_timer_gtky3_3 > 10
         & s1_timerpre1_3 > 10
         & s1_timerpre2_3 > 10
         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
         & s1_timerpost1_3 > 10
         & s1_timerpost2_3 > 10
         & s1_timerdemo1_3 > 10
         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

## Data selection

From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis. Units of study were excluded because the instructor did not participate in the study (33,483); the instructor or course wasn't actually eligible (96); the instructor administered the wrong survey (31); the student did not consent  (116); the student did not complete the first survey (304); the instructor did not complete the second survey (107); or one participant did not spend more than a second reading the consent page and more than ten seconds reading each page with five questions or more (428). The time limits are somewhat arbitrary, and different choices could have different analysis results.

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many students were discarded because the instructor did not participate? 33175
all_students <- isrdata %>%
  unique() %>%
  count() %>%
  pull()

all_students - isrdata %>%
  filter(!is.na(teacherid)) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many were excluded because the course wasn't actually eligible? 96
instructor_participated <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

instructor_participated - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many were excluded because the instructor didn't adminster the correct survey? 31
eligible <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

eligible - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors did not complete survey 1? 0
eligible_correct <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

eligible_correct - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many students did not consent? 116
teacher_complete <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

teacher_complete - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many students did not complete survey 1? 304
student_consent <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

student_consent - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1
         & s1_fullsurvey == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many were discarded because instructors didn't complete the second survey? 107
student_complete <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1
         & s1_fullsurvey == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

student_complete - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1
         & s1_fullsurvey == 1
         & t2_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many were discarded because someone didn't spend enough time? 428
instructor_complete <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1
         & s1_fullsurvey == 1
         & t2_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

instructor_complete - isrdata %>%
  filter(!is.na(teacherid)
         & t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & t1_timer_consent_3 > 1
         & s1_timer_consent_3 > 1
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & t2_timer > 10
         & s1_timer_gtky1_3 > 10
         & s1_timer_gtky2_3 > 10
         & s1_timer_gtky3_3 > 10
         & s1_timerpre1_3 > 10
         & s1_timerpre2_3 > 10
         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
         & s1_timerpost1_3 > 10
         & s1_timerpost2_3 > 10
         & s1_timerdemo1_3 > 10
         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

# Models

## Replication models

$treatment_{i}$ is the indicator that treatment was given.

$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study). Different covariates are used to model different outcomes. The original study used gender and CGPA as controls for grade based outcomes only. @citeoriginal state this is because females earn higher grades generally. This paper uses them as controls for the majority of outcomes. This is because of the potential link between student performance and perceptions of similarity and ISR. ISR anticipation is used as a control only for end-of-term student ISR perception.

$X_{2j}$ is a vector of instructor-level covariates (class size). The original study considered instructor fixed effects instead. Because each participating instructor taught one class in the study, the class sizes correspond directly to the instructor IDs, so it was inadvisable to use both. This paper selected class size because of the larger potential for transferable insights. These particular instructors are not of general interest, but the effect of class size may be.

$\epsilon_{ij}$ is a standard error. As this paper does not use clustered standard errors like @citeoriginal, there are slight differences in standard errors and p-values when compared with the original paper. This does not affect the regression coefficients.

In the resulting models, $\beta_{0}$ and $a_{k}$ are constant terms. $\beta_{1}$ is the coefficient that indicates treatment effect. $\Gamma_{1}$ and $\Gamma_{2}$ are coefficients for the covariates.

### Linear models

Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes. This includes complete scale outcomes: initial student similarity perception, end-of-term student similarity perception, end-of-term student ISR perception, and end-of-term instructor ISR perception. Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, and they can be treated as continuous variables and modeled using linear functions. It also includes grade-based outcomes: course grade and objectively graded exam grade, both of which are shown on a 4.0 GPA scale.

\begin{equation}
(\#eq:linmodel)
Outcome_{ij} = \beta_{0} + \beta_{1}treatment_{i} + \Gamma_{1}X_{1i} + \Gamma_{2}X_{2j} + \epsilon_{ij}
\end{equation}

### Ordinal logistic models

Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.The instructors were only asked to answer one five-point scale question about similarity, yielding an ordinal scale rather than a pseudo-continuous scale. Therefore, this outcome is treated differently and the model selected is specifically for ordinal outcomes, where the variable is ordered but not equally spaced. Because this variable is measured differently than the two student similarity perception scales, and therefore modeled differently, direct comparisons between student and instructor perception of similarity should be made with caution.

\begin{equation}
(\#eq:ordlogmodel)
prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + \Gamma_{1}X_{1i} + \Gamma_{2}X_{2j} + \epsilon_{ij} > k
\end{equation}

### Logistic models

Equation \@ref(eq:logmodel) is @citeoriginal's logistic model, used for the binary outcome, which is enrollment in Fall term 2017. Students are either not enrolled (0), enrolled (1), or they have graduated and are excluded from the model.

\begin{equation}
(\#eq:logmodel)
prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + \Gamma_{1}X_{1i} + \Gamma_{2}X_{2j} + \epsilon_{ij} > k
\end{equation}

## Additional models

When creating additional models, I focused on three key outcomes: standardized course grade, end-of-term student perceived ISR scale, and end-of-term student perceived similarity scale. I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance), and it is much more likely to detect an impact in the immediate course outcomes as opposed to outcomes in subsequent terms; and because previous studies focus more on student perception than instructor perception. Equation \@ref(eq:mylinmodel) shows the linear model used. $X_{3jk}$ represents student-instructor commonality covariates (matching racial self-ID, matching gender self-ID, age difference).

\begin{equation}
(\#eq:mylinmodel)
Outcome_{ij} = \beta_{0} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + X_{3jk}\Gamma_{3} + \epsilon_{ij}
\end{equation}

# Results

All results are reported using the treatment coefficient, $\beta_{1}$, and covariate coefficients, which are represented with $\Gamma$. 95% confidence intervals for those coefficients are shown in brackets below the coefficients. Significance levels are indicated with stars. Initial perceptions are indicated with 1, and end-of-term perceptions with 2. ISR refers to instructor-student relationship perception, and sim. refers to similarity perception. Final grade refers to the grade given for an objectively graded final exam in the course, and Std. course grade refers to a standardized final course grade.

## Replication results

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all replication models
studentfirstsimilaritymodel <- lm(data = expdata,
                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondsimilaritymodel <- lm(data = expdata,
                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondrelationshipmodel <- lm(data = expdata,
                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondrelationshipmodel <- lm(data = expdata,
                                     formula = t2_tsr ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
grademodel <- lm(data = expdata,
                 formula = grade ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
                     formula = t2_finalexam ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                data = expdata,
                                Hess = TRUE,
                                     na.action = na.omit)
enrollmentfallmodel <- glm(data = expdata,
                           formula = f17_enrolled ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                           family = "binomial",
                                     na.action = na.omit)
```

In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta_{1}$ = 0.194, 95% CI [0.136, 0.253]). Similarly, Model 2 shows that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta_{1}$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller. Both coefficients are minimal on a scale of 1 to 5. The adjusted R-squared values, which indicate the proportion of variation in the outcome that can be explained by the treatment and covariates, are low (0.032 and 0.022). These models are helpful for analyzing treatment effect, but they do not aim to be and are not useful predictors of student similarity perception, not even within the sample.

Model 3 shows no evidence that treatment impacted student end-of-term ISR perception. However, it does show a significant association between student ISR anticipation and student end-of-term ISR perception ($\Gamma$ = 0.671, 95% CI [0.624, 0.717]). For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.671. Models 4, 5, 6, 7, and 8 show no significant treatment effect on end-of-term instructor ISR perception, course grade, objectively graded exam grade, end-of-term instructor similarity perception, or enrollment in the subsequent term.

All models in Table \@ref(tab:modeltable) use gender, prior GPA, and year of study as student-level covariates. Instructor-level covariates are limited to the class size. This is based on the assumption that student traits will be more likely to affect student experiences, and therefore student outcomes. Race/ethnicity is not included because that covariate is further explored in separate models.

Model performance and validation for the two models with significant treatment effects, Models 1 and 2 from Table \@ref(tab:modeltable), can be found in [Appendix D].

```{r modeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodel,
          studentsecondsimilaritymodel,
          studentsecondrelationshipmodel,
          teachersecondrelationshipmodel,
          grademodel,
          examgrademodel,
          teachersecondsimilarity,
          enrollmentfallmodel,
          title = "Replication model results",
          align = TRUE,
          ci = TRUE,
          float.env = "sidewaystable",
          notes.label = "",
          font.size = "small",
          no.space = TRUE,
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-10pt",
          label = "tab:modeltable",
          type = "latex",
          header = FALSE,
          initial.zero = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Instructor sim. 2",
                             "Persistence"),
          covariate.labels = c("Treatment", "Student ISR 1", "Class size", "Student female", "CGPA", "Year of study"))
```

Because some students are more commonly targeted by persistence interventions, the subset of Hispanic and/or Black college student participants is considered separately from the rest of the student population. While the effect of treatment on student similarity perception is larger than for the general student population ($\beta_{1}$ = 0.199, 95% CI [0.124, 0.274] for initial perception; $\beta_{1}$ = 0.116, 95% CI [0.021, 0.211] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes. Full results are in [Appendix E].

Similarly, the subset of first-generation college student participants is considered. The effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta_{1}$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta_{1}$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), but there is still no effect on ISR perception, instructor perception, or student outcomes. Full results are in [Appendix F].

### Exploratory replication results

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all replication models

studentfirstisrmodel1g <- lm(data = expdata,
                                  formula = s1_tsr ~ s_firstgen + s1_female + ir_f16_gpa + grade + year + n_course,
                                     na.action = na.omit)
studentsecondrelationshipmodel1g <- lm(data = expdata,
                                     formula = s2_tsr ~ s_firstgen + s1_female + ir_f16_gpa + grade + year + n_course,
                                     na.action = na.omit)
teachersecondrelationshipmodel1g <- lm(data = expdata,
                                     formula = t2_tsr ~ s_firstgen + s1_female + ir_f16_gpa + grade + year + n_course,
                                     na.action = na.omit)
```

Table \@ref(tab:modeltablerel1) disregards the treatment condition, instead considering the association between student identity and similarity/ISR. While we cannot make causal inferences, this is an exploratory analysis that could shape future intervention studies. If a student identifies as first generation, can we anticipate a different student or instructor ISR perception? In this case, there is no significant association.

```{r modeltablerel1, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstisrmodel1g,
          studentsecondrelationshipmodel1g,
          teachersecondrelationshipmodel1g,
          title = "Replication model results: first-generation status and ISR",
          align = TRUE,
          ci = TRUE,
          no.space = TRUE,
          initial.zero = FALSE,
          notes.label = "",
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-10pt",
          label = "tab:modeltablerel1",
          font.size = "small",
          type = "latex",
          header = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student ISR 1",
                             "Student ISR 2",
                             "Instructor ISR 2"),
          covariate.labels = c("Student first gen.", "Student female", "CGPA", "Course grade", "Year of study", "Class size"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all replication models
studentfirstisrmodelhb <- lm(data = expdata,
                                  formula = s1_tsr ~ s_hisp_black + s1_female + ir_f16_gpa + grade + year + n_course,
                                     na.action = na.omit)
studentsecondrelationshipmodelhb <- lm(data = expdata,
                                     formula = s2_tsr ~ s_hisp_black + s1_female + ir_f16_gpa + grade + year + n_course,
                                     na.action = na.omit)
teachersecondrelationshipmodelhb <- lm(data = expdata,
                                     formula = t2_tsr ~ s_hisp_black + s1_female + ir_f16_gpa + grade + year + n_course,
                                     na.action = na.omit)

```

Table \@ref(tab:modeltablerel2), however, shows that there is a significant negative association between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\Gamma$ = -0.096, 95% CI [-0.166, -0.026]).

```{r modeltablerel2, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstisrmodelhb,
          studentsecondrelationshipmodelhb,
          teachersecondrelationshipmodelhb,
          title = "Replication model results: Hispanic and/or Black students and ISR",
          align = TRUE,
          ci = TRUE,
          no.space = TRUE,
          initial.zero = FALSE,
          notes.label = "",
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-10pt",
          label = "tab:modeltablerel2",
          font.size = "small",
          type = "latex",
          header = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student ISR 1",
                             "Student ISR 2",
                             "Instructor ISR 2"),
          covariate.labels = c("Student Hispanic and/or Black", "Student female", "CGPA", "Course grade", "Year of study", "Class size"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all replication models
studentsecondrelationshipmodel_ <- lm(data = expdata,
                                     formula = s2_tsr ~ treatment + s2_sim + t2_sim1 + n_course,
                                     na.action = na.omit)
teachersecondrelationshipmodel_ <- lm(data = expdata,
                                     formula = t2_tsr ~ treatment + s2_sim + t2_sim1 + n_course,
                                     na.action = na.omit)
grademodel_ <- lm(data = expdata,
                 formula = grade ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                                     na.action = na.omit)
examgrademodel_ <- lm(data = expdata %>% filter(obj_exam == 1),
                     formula = t2_finalexam ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                                     na.action = na.omit)
enrollmentfallmodel_ <- glm(data = expdata,
                           formula = f17_enrolled ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                           family = "binomial",
                                     na.action = na.omit)
```

Table \@ref(tab:noncausaltable) considers the association between similarity perception and ISR perception and the relationship between ISR perception and student outcomes. Again, because only the treatment was randomized, we cannot make causal inferences. Course size is used as a covariate for ISR perception because it directly shapes how students and instructors relate to one another. Student gender and prior GPA are used as covariates for student outcomes for the reasons previously outlined from @citeoriginal.

In Models 1 and 2, we can see that student and instructor similarity perception are both significantly associated with student and instructor ISR perception. Unsurprisingly, student similarity perception is more strongly associated with student ISR perception ($\Gamma$ = 0.733, 95% CI [0.706, 0.76]), and instructor similarity perception is more strongly associated with instructor ISR perception ($\Gamma$ = 0.512, 95% CI [0.483, 0.541]).

Models 3 and 4 show that student ISR perception is significantly associated with course grade, while instructor ISR perception is significantly associated with course grade and final grade. Instructor perception is more strongly associated ($\Gamma$ = 0.34, 95% CI [0.295, 0.385] for course grade; $\Gamma$ = 0.643, 95% CI [0.642, 0.759] for final grade) than student perception ($\Gamma$ = 0.115, 95% CI [0.064, 0.164] for course grade).

Model 5 indicates that ISR perception is not significantly associated with persistence.

```{r noncausaltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentsecondrelationshipmodel_,
          teachersecondrelationshipmodel_,
          grademodel_,
          examgrademodel_,
          enrollmentfallmodel_,
          title = "Replication model results: Similarity, ISR, and student outcomes",
          align = TRUE,
          ci = TRUE,
          initial.zero = FALSE,
          no.space = TRUE,
          font.size = "small",
          notes.label = "",
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-10pt",
          label = "tab:noncausaltable",
          type = "latex",
          header = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Persistence"),
          covariate.labels = c("Treatment",
                             "Student sim. 2",
                             "Instructor sim. 2",
                             "Class size",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Student female",
                             "CGPA"))
```

## Additional results

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# create variables to indicate matching student-instructor traits
expdata <- expdata %>%
  mutate(race_match = ifelse(t1_race_1 == s1_race_1
                             & t1_race_2 == s1_race_2 
                             & t1_race_3 == s1_race_3 
                             & t1_race_4 == s1_race_4 
                             & t1_race_5 == s1_race_5 
                             & t1_race_6 == s1_race_6 
                             & t1_race_7 == s1_race_7,
                             1,
                             0)) %>%
  mutate(gend_match = ifelse(t_female == s1_female,
                             1,
                             0)) %>%
  mutate(age_dif = t_age - s1_age)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all models for matching traits
mystudentsecondsimilaritymodel <- lm(data = expdata,
                                     formula = s2_sim ~ treatment + n_course + race_match + gend_match + age_dif + nsims,
                                     na.action = na.omit)
mystudentsecondrelationshipmodel <- lm(data = expdata,
                                     formula = s2_tsr ~ treatment + n_course + race_match + gend_match + age_dif + nsims,
                                     na.action = na.omit)
mystdgrademodel <- lm(data = expdata,
                      formula = grade ~ treatment + ir_f16_gpa + n_course + race_match + gend_match + age_dif + nsims,
                                     na.action = na.omit)
```

Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits. As with the replication results, there is a measurable treatment effect for initial student similarity perception ($\beta_{1}$ = 0.108, 95% CI [0.036, 0.179]). However, the coefficient is even lower than in the replication models. There is still no treatment effect for student end-of-term ISR perception or standardized course grade. Because the matching ID covariates were used, student covariates like gender and race were not included. Prior GPA and class size were still included, because they do not have a "matching" trait.

```{r mymodeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(mystudentsecondsimilaritymodel,
          mystudentsecondrelationshipmodel,
          mystdgrademodel,
          title = "Additional model results",
          align = TRUE,
          ci = TRUE,
          no.space = TRUE,
          notes.label = "",
          initial.zero = FALSE,
          font.size = "small",
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-10pt",
          label = "tab:mymodeltable",
          type = "latex",
          header = FALSE,
          dep.var.labels = c("Student sim. 2",
                             "Student ISR 2",
                             "Std. course grade"),
          covariate.labels = c("Treatment",
                               "CGPA", 
                               "Class size",
                               "Matching racial self-ID", "Matching gender self-ID", "Age difference", "Number of similarities")
          )
```

# Discussion

The results of this replication study align with @citeoriginal's findings. When undergraduate students and their instructors learn about similarities they share, there is a small improvement in student similarity perception. However, the treatment has no significant impact on student ISR perception, instructor perception, or student outcomes.

To understand what was and wasn't impacted by the treatment, we can consider the desired causal sequence in this study, shown in Figure \@ref(fig:causaldiagram). The intervention refers to informing students and instructors about shared traits. It does not target actual instructor-student similarity -- it simply aims to heighten awareness of any existing similarities. Similarity perception is measured by the two perceived similarity scales. ISR, or instructor-student relationship, cannot be directly measured -- only perception can be measured, through the perceived ISR scales. The student outcomes, grades and persistence, are more easily measured. The diagram shows the complexity of succeeding with this type of intervention: the treatment must affect perception, which must then affect relationships, which must then affect student outcomes.

```{r causaldiagram, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Diagram of desired causal sequence", dev='png', out.width="100%"}
# Make a diagram to show the causal sequence that is desired in this intervention
DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = LR]

  node [shape = diamond, color = MediumSpringGreen, penwidth = 2]         
  treatment [label = 'Intervention']

  node [shape = rectangle, color = MediumSlateBlue]        
  ssp [label = 'Student\nSimilarity\nPerception']
  isp [label = 'Instructor\nSimilarity\nPerception']
  isr [label =  'ISR']

  node [shape = oval, color = HotPink]
  grades [label = 'Grades']
  persistence [label = 'Persistence']

  # edge definitions with the node IDs
  treatment -> ssp [ label = '1']
  treatment -> isp [ label = '2']
  ssp -> isr [ label = '3']
  isp -> isr [ label = '4']
  isr -> grades [ label = '5']
  isr -> persistence [ label = '6']
  grades -> persistence [ label = '7']
  }",
  height = 564,
  width = 2700)
```

This study did successfully demonstrate relationship 1: the treatment caused a statistically significant increase in student similarity perception at the beginning and end of the term. Relationship 2 failed to materialize, as there was no significant difference in instructor similarity perception in the treatment group. To understand why the treatment effect was small for students and insignificant for instructors, we can consider the specifics of the treatment. What similarities were students and instructors informed of? [Appendix G] contains a complete list. Students and instructors were presented with generalities such as "An instructor’s most important characteristic? You both say being knowledgeable" or "You are cut from the same cloth–you both described yourself as thoughtful!" Such facts may not be unusual enough to be memorable, reducing the treatment's impact. It is also possible that similarities declared by an anonymous third party are simply not as impactful as similarities that are personally observed and experienced. Finally, because the instructor similarity perception was measured on one five-point scale, rather than a full scale as student similarity perception was, perhaps the measurement was less sensitive to change. @citeoriginal note that they used a single question to reduce the burden placed on instructors, who had to answer surveys for many students. However, this courtesy created a dissimilarity in measurment for outcomes that we would like to compare.

The exploratory modeling in Table \@ref(tab:noncausaltable) indicates that relationships 3 and 4 may exist: increases in similarity perception were associated with statistically significant increases in ISR perception. However, this relationship was not the focus of the study, and we cannot draw any firm conclusions. We also cannot determine causality: it is possible that an improved instructor-student relationship results in an increase in perceived similarity, not vice-versa. Additionally, much of the research about similarity and ISRs comes from elementary and secondary school contexts. The exact relationship between perceived similarity and ISRs could be different at the college level.

The exploratory modeling in Table \@ref(tab:noncausaltable) also indicates that ISR, or at least ISR perception, is related to student grade outcomes, as in relationship 5 in Figure \@ref(fig:causaldiagram). Again, we cannot conclude anything about causality. It is possible that students engage in behaviors that both earn higher grades and build better relationships with their instructors. Because the quality of the ISR cannot be measured directly, it is also possible that instructor-student relationships can improve student outcomes, but the particular perceptions that are measured in this case are not the relevant factors, or perceived ISR is less important that the real behaviors within an ISR. Even in this exploratory analysis, ISR perception was not associated with persistence, and relationship 6 did not appear.

The environment of @citeoriginal's experiment is relevant. They aimed to build ISRs in a context where a student's average class size is 49.5, students typically take five classes at a time, and instructors teach multiple classes, sometimes at multiple campuses. The intervention is appealingly simple, quick, and inexpensive, but such a small change is likely insignificant in the face of a wider campus culture that makes connections challenging.

Further, the study indicates that the instructor and student populations are demographically different. Compared to the group of instructors, the student body had 3.5 times as many individuals who identified as Hispanic. Compared to students, the instructor population had 3.1 times as many individuals who identified as White. This is only a hint of the differences that may exist between the student and instructor populations. As most students will not go on to be instructors, we can expect, at the very least, that many students will not share life goals and trajectories with their instructors. If students feel fundamentally different from their instructors, based on life experience or socio-economic status, an intervention that provides superficial similarities is insufficient to bridge that gap.

If colleges are considering investing in interventions that try to enhance similarity perception, we need to ask what it means to perceive similarity. Do some similarities matter more than others in ISRs? As indicated by the results in Table \@ref(tab:modeltablerel2), some marginalized student identities may be associated with measurably lower instructor ISR perceptions. If demographics traits, which are already part of an inequitable campus environment, are important for perceived similarity, should positive ISRs be contingent on similarity at all? There are other ways to cultivate ISRs that may make more sense and offer more equitable outcomes.

## Limitations

The study's internal validity comes from the successfully randomly-assigned treatment and control groups, but there is one concern. The instructors who participated were part of both the treatment and control group. That is, they received similarities for some students but not others. If more significant effect sizes, or larger effect sizes, had been found, this potential lack of separation between the treatment and control groups would be a problem, because the instructor's attention could have simply been shifted from students in the control group to students in the treatment group, potentially exaggerating the effect. However, as little treatment effect was found, this limitation is less concerning.

The external validity is more questionable. The study participants were chosen in two main ways: the instructor was interested in the study and completed all the steps successfully, and the student agreed to participate. Data is only available for the second student survey for students who chose to follow through. The sample of instructors that would be interested in the study would not necessarily be comparable to the overall population: they could be instructors who care more about supporting the work of other academics, instructors for whom a $150 giftcard was a significant draw, or instructors who are particularly interested in the role of ISRs. Because the instructors administered the surveys to students, students who actively engage with their instructors' requests could be overrepresented in the student sample.

Any measurable effects could also be specific to this particular college campus. Interventions may not be transferable due to different demographics, cultures, and institutional structures.

As with all survey studies, there is a risk that self-reported information is inaccurate. In particular, the surveys used by @citeoriginal had many Likert scale questions about very similar topics. Survey fatigue may have reduced the meaningfulness of the survey answers, especially those later in the survey. While responses that were given overly quickly were removed from the analysis, the fact that 428 cases had to be removed for this reason indicates a common problem with survey data: we cannot know if the participants actually considered their answers. Additionally, survey-takers may feel pressure to give the answers they believe they should give, rather than answers that are more honest. This could be especially true in this context, where students are usually being graded on their performance and answers to multiple-choice questions.

## Future directions

The data collected by @citeoriginal is wider than is shown in this paper or the original. As mentioned previously, there were other survey questions targeting interest level, enjoyment, and confidence, amoungst other additional information. This data could be used in further exploratory analysis to look at associations between similarity, ISR, interest, enjoyment, and confidence.

Given the minimal treatment effects found in this particular intervention, future studies could take a wide variety of approaches. First, more intensive, but similar, interventions to increase perceived similarity could be tested. These interventions could include more specific or important similarities between students and instructors. Second, different methods of increasing perceived similarity, perhaps based in actual instructor-student interaction, could be tested. Such interventions are less desirable from the college's point of view, because they would be more costly and demanding to implement. Third, interventions that aim to improve ISR without increasing perceived similarity could be considered. Similarity is only one of many theorized methods for improved ISRs.

\newpage

# (APPENDIX) Appendix {-}

# Appendix A

Complete list of key measures

* Initial student survey
  1. Student perceived similarity scale
      - Overall, how similar to your instructor's values do you think your values are?
      - How similar are your goals for the course and your instructor's goals?
      - In general, how similar do you think your views about the course content and your instructor's are?
      - How much do you think you have in common with your instructor?
      - How similar do you think your personality is compared to your instructor's?
      - Overall, how similar do you think you and your instructor are?
  2. Student anticipated ISR scale
      - How much do you think you will enjoy learning from this instructor?
      - How friendly do you think this instructor will be towards you?
      - How encouraging do you think this instructor will be towards you?
      - If you came back to visit this instructor three years from now, how excited do you think they would be?
      - How motivating do you think you will find this instructor's class?
      - How caring do you think this instructor will be towards you?
      - Overall, how much do you think you will learn from this instructor?
* End of term student survey
  3. Student perceived similarity scale
  4. Student perceived ISR scale
      - How much do you enjoy learning from this professor?
      - How friendly do you think this professor is towards you?
      - If you came back to visit this professor three years from now, how excited do you think they would be?
      - How motivating do you find this professor's class?
      - How caring do you think this professor is towards you?
      - How encouraging do you think this professor is towards you?
      - Overall, how much do you think you have learned from this professor?
* End of term instructor survey
  5. Instructor similarity perception
      - Overall, how similar do you think you and STUDENTNAME are?
  6. Instructor perceived ISR scale
      - How much did you enjoy helping STUDENTNAME learn?
      - How caring was STUDENTNAME towards you?
      - How often did you say something encouraging to STUDENTNAME?
      - How friendly was STUDENTNAME towards you?
      - If this student came back to visit you three years from now, how excited would you be?
      - How motivating did STUDENTNAME find the activities that you plan for class?
      - Overall, how much did STUDENTNAME learn from you?
  8. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
* University internal records:
  9. Course grade: The final grade that the student received in the course.
  10. Standardized course grade: The student's final grade, standardized against other grades in the course.
  11. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.

\newpage

# Appendix B

```{r demotable3, echo=FALSE, warning=FALSE, message=FALSE}
# Make a summary table of nominal variables, by treatment group.
demotab3 <- expdata %>% 
  dplyr::select(treatment, 
                t_female, 
                t_race, 
                t_firstgen,
                t_age,
                n_course) %>%
  mutate(treatment = ff_label(treatment, "Group"),
         t_female = ff_label(t_female, "Instructor gender"), 
         t_age = ff_label(t_age, "Instructor age"),
         t_race = ff_label(t_race, "Instructor race/ethnicity"), 
         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status"), 
         n_course = ff_label(n_course, "Class size")) %>%
  summary_factorlist("treatment", 
                     c("t_female", "t_race", "t_firstgen", "t_age", "n_course"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N(%) = ") %>%
  rename(N = "Total N", Missing = "Missing N")

# Replace variables that are encoded with their true meanings
demotab3[2:3,4] <- c("Male", "Female")
demotab3[10:11,4] <- c("No", "Yes")

# Put it all in a nice table.
demotab3 %>%
  knitr::kable(caption = "Teacher covariates for treatment and control groups",
               booktabs = TRUE, linesep = "",
               ) %>%
  column_spec(5, width = "5em") %>%
  column_spec(6, width = "5em") %>%
  column_spec(7, width = "5em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped", "hold_position"),
                            stripe_index =c(2:3, 10:11, 13),
                            stripe_color = "#f6f0fb")
```

\newpage

# Appendix C

```{r missingdata, echo=FALSE, warning=FALSE, message=FALSE, fig.height=9, fig.width=9}
# Plot missing data patterns for key covariates
explanatory = c("s2_tsr", "s1_tsr", "ir_f16_gpa", "n_course", "s1_female")
dependent = "treatment"

expdata %>%
  dplyr::select(treatment, 
                s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, 
                grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, 
                s1_female, n_course) %>%
  missing_pairs(dependent, explanatory)
```

\newpage

# Appendix D

Linear models are based on a number of assumptions: a linear relationship, homoscedasticity (homogeneity of variance), a lack of correlation between the independent variables, and normally-distributed residuals. Figure  \@ref(fig:model1performance) summarizes tests of these assumptions for the initial student similarity perception model. All assumptions seem to hold for this model. Figure  \@ref(fig:model2performance) shows the same tests for the end-of-term student similarity perception model.

```{r model1performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="Initial student similarity perception model assumptions"}
# Plot Student sim. 1 model assumptions
performance::check_model(studentfirstsimilaritymodel)
```

```{r model2performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="End-of-term student similarity perception model assumptions"}
# Plot student sim. 2 model assumptions
performance::check_model(studentsecondsimilaritymodel)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# Split train and test data
set.seed(1847)
expdata_split <- rsample::initial_split(expdata, prop = 0.80)
expdata_train <- rsample::training(expdata_split)
expdata_test  <-  rsample::testing(expdata_split)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all replication models
studentfirstsimilaritymodelv <- lm(data = expdata_train,
                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondsimilaritymodelv <- lm(data = expdata_train,
                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
```

Table \@ref(tab:modelvtable) shows the two primary significant models, redone using a random train/test split (80% training, 20% testing). The treatment effect coefficient for initial student similarity perception is close to the coefficient in the first version of the model ($\beta_{1}$ = .202, 95% CI [.137, .267]), but the coefficient for end-of-term student similarity perception differs more ($\beta_{1}$ = .137, 95% CI [.055,.219]). Table \@ref(tab:traintesttable) shows metrics to assess the performance of the two models on the training and testing datasets. The root mean squared errors and mean absolute errors are slightly larger for the testing data, which makes sense, as the model was fit to the training data but not the testing data. From these numbers, we can conclude that minimal overfitting in the model has occurred.

```{r modelvtable, results=c('asis', 'hold'), echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodelv,
          studentsecondsimilaritymodelv,
          title = "Train replication model results",
          align = TRUE,
          ci = TRUE,
          no.space = TRUE,
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-10pt",
          label = "tab:modelvtable",
          font.size = "small",
          notes.label = "",
          type = "latex",
          header = FALSE,
          initial.zero = FALSE,
          keep = c("treatment", "n_course", "s1_female", "ir_f16_gpa", "year", "Constant"),
          #omit = c("teacherid"),
          #mit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2"),
          covariate.labels = c("Treatment", "Class size", "Student female", "CGPA", "Year of study"))
```

```{r traintesttable, echo=FALSE, warning=FALSE, message=FALSE}
# Use the model to predict training data outcomes
expdata_train <- expdata_train %>%
  mutate(pred1 = predict(studentfirstsimilaritymodelv, expdata_train),
         pred2 = predict(studentsecondsimilaritymodelv, expdata_train))

# Use the model to predict testing data outcomes
expdata_test <- expdata_test %>%
  mutate(pred1 = predict(studentfirstsimilaritymodelv, expdata_test),
         pred2 = predict(studentsecondsimilaritymodelv, expdata_test))

# Calculate metrics to assess model performance
metrics(expdata_train, s1_sim, pred1) %>%
  bind_cols(metrics(expdata_test, s1_sim, pred1) %>% dplyr::select(.estimate)) %>%
  bind_rows(metrics(expdata_train, s2_sim, pred2) %>%
              bind_cols(metrics(expdata_test, s2_sim, pred2) %>% dplyr::select(.estimate))) %>%
  mutate(.metric = c("Root mean square error", "R-squared", "Mean absolute error",
                     "Root mean square error", "R-squared", "Mean absolute error")) %>%
# put in a table
kable(caption = "Metrics for student similarity perception models",
        booktabs = TRUE,
        col.names = c("Metric", "Estimator", "Training data", "Testing data")) %>%
  pack_rows("Initial", 1, 3) %>%
  pack_rows("End-of-term", 4, 6) %>%
  kable_styling(latex_options = "hold_position")
```

\newpage

# Appendix E

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Filter to data that only includes students who identify as hispanic or black
expdatahb <- expdata %>%
  filter(s_hisp_black == 1)

# Make all replication models
studentfirstsimilaritymodelhb <- lm(data = expdatahb,
                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondsimilaritymodelhb <- lm(data = expdatahb,
                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondrelationshipmodelhb <- lm(data = expdatahb,
                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondrelationshipmodelhb <- lm(data = expdatahb,
                                     formula = t2_tsr ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
grademodelhb <- lm(data = expdatahb,
                 formula = grade ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
examgrademodelhb <- lm(data = expdatahb %>% filter(obj_exam == 1),
                     formula = t2_finalexam ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                data = expdatahb,
                                Hess = TRUE,
                                     na.action = na.omit)
enrollmentfallmodelhb <- glm(data = expdatahb,
                           formula = f17_enrolled ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                           family = "binomial",
                                     na.action = na.omit)
```

```{r modeltablehb, results=c('asis','hold'), echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodelhb,
          studentsecondsimilaritymodelhb,
          studentsecondrelationshipmodelhb,
          teachersecondrelationshipmodelhb,
          grademodelhb,
          examgrademodelhb,
          teachersecondsimilarityhb,
          enrollmentfallmodelhb,
          title = "Replication model results: Hispanic and Black students only",
          align = TRUE,
          ci = TRUE,
          initial.zero = FALSE,
          float.env = "sidewaystable",
          no.space = TRUE,
          notes.label = "",
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-12pt",
          label = "tab:modeltablehb",
          font.size = "small",
          type = "latex",
          header = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Instructor sim. 2",
                             "Persistence"),
          covariate.labels = c("Treatment", "Student ISR 1", "Class size", "Student female", "CGPA", "Year of study"))
```

\newpage

# Appendix F

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Filter so we only have data for first-generation college students
expdata1g <- expdata %>%
  filter(s_firstgen == 1)

# Make all replication models
studentfirstsimilaritymodel1g <- lm(data = expdata1g,
                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondsimilaritymodel1g <- lm(data = expdata1g,
                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondrelationshipmodel1g <- lm(data = expdata1g,
                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondrelationshipmodel1g <- lm(data = expdata1g,
                                     formula = t2_tsr ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
grademodel1g <- lm(data = expdata1g,
                 formula = grade ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
examgrademodel1g <- lm(data = expdata1g %>% filter(obj_exam == 1),
                     formula = t2_finalexam ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                data = expdata1g,
                                Hess = TRUE,
                                     na.action = na.omit)
enrollmentfallmodel1g <- glm(data = expdata1g,
                           formula = f17_enrolled ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                           family = "binomial",
                                     na.action = na.omit)
```

```{r modeltable1g, results=c('asis','hold'), echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodel1g,
          studentsecondsimilaritymodel1g,
          studentsecondrelationshipmodel1g,
          teachersecondrelationshipmodel1g,
          grademodel1g,
          examgrademodel1g,
          teachersecondsimilarity1g,
          enrollmentfallmodel1g,
          title = "Replication model results: first-generation students only",
          align = TRUE,
          initial.zero = FALSE,
          ci = TRUE,
          notes.label = "",
          float.env = "sidewaystable",
          no.space = TRUE,
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-12pt",
          font.size = "small",
          label = "tab:modeltable1g",
          type = "latex",
          header = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Instructor sim. 2",
                             "Persistence"),
          covariate.labels = c("Treatment", "Student ISR 1", "Class size", "Student female", "CGPA", "Year of study"))
```

\newpage

# Appendix G

```{r similarities, echo=FALSE, warning=FALSE, message=FALSE}
# Collect all similarities that were presented to students and instructors
tibble(Similarities = unique(expdata$sentence1)) %>%
  bind_rows(tibble(Similarities = unique(expdata$sentence2))) %>%
  bind_rows(tibble(Similarities = unique(expdata$sentence3))) %>%
  bind_rows(tibble(Similarities = unique(expdata$sentence4))) %>%
  bind_rows(tibble(Similarities = unique(expdata$sentence5))) %>%
  bind_rows(tibble(Similarities = unique(expdata$sentence6))) %>%
  bind_rows(tibble(Similarities = unique(expdata$sentence7))) %>%
  unique() %>%
  arrange(Similarities) %>%
  filter(Similarities != " ") %>%
# put in a table
  knitr::kable(caption = "Similarities presented to students and instructors during treatment",
               booktabs = TRUE, linesep = "", longtable = TRUE
               ) %>%
  column_spec(1, width = "47em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
                            stripe_color = "#f6f0fb")
```

\newpage

# References